---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```


# Question 1.
Find the weather forecast for the next three days online. Create a list of lists in which each inner list should be labelled with the name of the day of the week. Each inner list should contain named elements telling you the temperature, chance of rain and wind speed.

```{r}

forecast <- list("Thursday" = list(temp = 21, rain = 0, wind = 5),
                 "Friday" = list(temp = 19, rain = 0.3, wind = 7),
                 "Saturday" = list(temp = 21, rain = 0.7, wind = 15))

```



# Question 2.
Using the list you’ve just made find:

The chance of rain tomorrow
The wind speed on the day after tomorrow
The temperature tomorrow.

```{r}

forecast$Thursday$rain

forecast[["Friday"]][["wind"]]

forecast[["Thursday"]][["temp"]]

#either method works fine here for exploring these lists
```



# Question 3.
Find a 14 day forecast online and create a matrix of the forecasts for temperature, where each week is one row and columns correspond to days.

You will want to give a name to the matrix you create.

```{r}

days_of_week <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")

fourteen_day_temp <- c(20, 21, 22, 18, 16, 13, 16, 20, 23, 18, 15, 14, 15, 16)

weather_matrix <- matrix(fourteen_day_temp, nrow = 2, ncol = 7)

rownames(weather_matrix) <- c("week1", "week2")
colnames(weather_matrix) <- days_of_week

weather_matrix
```



# Question 4.
From the matrix you just made select:

The third day of the second week
The whole second week
The first, third and fifth days in both weeks.

```{r}

weather_matrix["week2", "Wednesday"]  # can call these by row and column names OR
weather_matrix[2, ]                   # 2 works just as well as the row name
weather_matrix[ ,c(1,3,5)]

```



# Question 5.
Write a function that pre-appends “The answer is:” to any character input, while for numeric input, it either just returns the number for numbers less than 10, or divides the number by 10 for numbers greater than 10. For anything else return “Invalid input”.

```{r}

# look at all the lovely rainbow bracket colours


too_many_ifs <- function(x){
  if(is.character(x)){
    paste("The answer is:", x)
  } else {
    if(is.numeric(x) & x<10){
      x
    }  else {      
      if (is.numeric(x) & x>10) {
        x/10
      }else{
        print("Invalid Input")
      }
    }
  }
}

too_many_ifs("Neil")

too_many_ifs(2)

too_many_ifs(22)

# too_many_ifs(list("a"))


```

# Question 6.
Write a function which checks if two vectors have the same elements.

If they have the same elements in the same order, return “Exactly the same”
If they have the same elements but in a different order, return “The same”
If they don’t have the same elements, return “Different”.

```{r}

vector_check <- function(x, y){
  # lets sort for our second check here to make code more readable
  v <- sort(x)
  w <- sort(y)
  
  if(identical(x,y)){ # can use all(x == y) here instead
    print("Exactly the same")
  } else {
    if(identical(v,w)){ # can use all(x == y) here instead
      print("The same")
    } else{
      print("Different")
    }
  }
}

vector_check(c("a", "b"), c("a", "b"))

vector_check(c("b", "a"), c("a", "b"))

vector_check(c("b", "z"), c("a", "c"))

```


2 Extension


Question 7.
Write code that prints a multiplication table for numbers up to 12. i.e. there will be 12 rows and 12 columns, and each element should be the row number multiplied by the column number.

```{r}

multiplication_table <- matrix(nrow = 12, ncol = 12)

nrows <- nrow(multiplication_table)
ncols <- ncol(multiplication_table)

for (row_i in 1:nrows) {
 for (col_i in 1:ncols) {
   multiplication_table[row_i, col_i] <-  row_i*col_i
 }
}
multiplication_table
```




# Question 8.
Write a function that takes a vector, and creates a text art histogram, like the one below.

You may want to use the function cat(). This is similar to print(), but prints directly, without vector markers. You can also include newlines "\n" when using cat().

For example, the vector c(1, 2, 3, 4, 5, 1, 3, 4, 5, 1, 3, 4, 4, 3, 3, 3, 3) should produce the histogram below - the histogram dislpays the number of times each number appears in the vector.

1 | ***
2 | *
3 | *******
4 | ****
5 | **


```{r}

a <- c(1, 2, 3, 4, 5, 1, 3, 4, 5, 1, 3, 4, 4, 3, 3, 3, 3)
b <- c(2, 4, 5, 6, 7, 8, 9, 9, 9, 1, 2, 3, 4, 5, 6, 7 ,7, 7, 7, 7, 7)
c <- c(1, 2, 3, 4, 5, 1, 3, 4, 5, 1, 3, 4, 4, 3, 3, 3, 3, 1, 2, 3, 4, 5, 1, 3, 4, 5, 1, 3, 4, 4, 3, 3, 3, 3, 1, 2, 3, 4, 5, 1, 3, 4, 5, 1, 3, 4, 4, 3, 3, 3, 3)
text_art <- function(x){
  stars <- ("")
  
  for (i in 1:max(x)){
    sum_of_number <- sum(x == i)
    stars <- ("")
    for (j in 1:sum_of_number){
      stars <- paste("*", stars)
    }
    cat(i, " | ", stars, "\n")
  }
}
text_art(a)

text_art(b)

text_art(c)

```
