---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(CodeClanData)
```


dplyr practice lab
You now have an opportunity to practice the skills you’ve learnt today in a mini lab task. For this you will be using data which contains income generated by US states from year 2002 to 2015. The data is available from CodeClanData in the dataset state_income_data.

MVP

Question 1.

Load in the data, and have a look at it. How many variables does it have? How many observations? What are the variable names? What type of variables does it have?

51 x 16 tibble


```{r}
state_income_data %>% 
  glimpse()

```

```{r}
state_income_data %>% 
  summary()
```

```{r}
state_income_data %>% 
  view()
```

Question 2.

Select the columns which contain the information from years 2002 to 2008. Ensure you keep the Index and State variables as well.

```{r}
state_income_data %>% 
  select(Index, State, Y2002:Y2008)
```



Question 3.

Rename the Index column to first_letter. Hint: look up the rename() function.
```{r}
state_income_data %>% 
  rename(first_letter = Index)

```


Question 4.

Find the seven states which start with an A,B, or C.

```{r}
abc_states <- state_income_data %>% 
  filter(Index %in% c("A", "B", "C"))
abc_states
```



Question 5.

Find which of these seven states have the greatest income in 2008.

```{r}
abc_states %>% 
  filter(Y2008 == max(Y2008))
```


Question 6.

Calculate the change in income for each state of these seven states (states starting with A,B, or C) between 2002 and 2008. Save the income change in a new column called income_change.

```{r}
abc_states %>% 
  mutate(income_change = (Y2008 - Y2002), .before = Y2002)
```


Question 7.

Find which of these seven states income change is greater than 500,000 (i.e. an increase of 500,000 or decrease of 500,000).


```{r}
abc_states %>% 
  mutate(income_change = (Y2008 - Y2002)) %>% 
  filter(income_change > 500000 | income_change < -500000)
```




Question 8.

Calculate the mean and median income change for all seven states.

```{r}
abc_states %>% 
  mutate(income_change = (Y2008 - Y2002)) %>% 
  summarise(mean(income_change), median(income_change))
```


Question 9.

Go back to the original dataset which contains all states. Write a pipe which selects the State and Y2008 variables, arranges the income from highest to lowest, filters all those states which have a income of greater than 1 million, and then finally chooses the top 10 wealthiest states.

All rows have greater income than 1000000 so filter is irrelevant?

```{r}
state_income_data %>% 
  select(State, Y2008) %>% 
  arrange(desc(Y2008)) %>% 
  filter(Y2008 > 1000000) %>% 
  slice_head(n = 10)



```

