---
title: "R Notebook"
output: html_notebook
---
```{r}
library(tidyverse)

tweets <- read_csv("data/code_clan_tweets.csv")
```

# Question 1.

Load the code_clan_tweets.csv data. Find the number of rows, columns, and list all the variable names.

```{r}
tweets
glimpse(tweets)

psych::describe(tweets)

skimr::skim(tweets)

colnames(tweets) # lists the variable names as per the question
```


# Question 2.

Find the total number of favourites (stored in favorite_count) that CodeClan tweets have got. Don’t add a tweet’s favorite_count to the total when the tweet was quoted (see the is_quote variable).

```{r}
tweets %>% 
  filter(is_quote == FALSE) %>% # can just use filter(is_quote) here as it is already a boolean vector
  summarise(sum(favorite_count))

# 425 favourites
```


# Question 3.

Summarise the mean number of retweets (retweet_count), split by the type of platform that was used to tweet (source). Again, omit quoted tweets from the mean.

```{r}
tweets %>% 
  filter(is_quote == FALSE) %>% 
  group_by(source) %>% 
  summarise(mean_retweets = round(mean(retweet_count), 2))
```



# Question 4.

Count the total number of likes (i.e. total of favorite_count), split by media type, and arrange them from most likes to least. Assume that any tweets without a listed media type are of type “text”.

```{r}

tweets %>% 
  count(media_type)
  # there are 145 NAs lets change these to "text"?
 
 tweets %>% 
   mutate(media_type = if_else(is.na(media_type), "text", media_type)) %>% 
   group_by(media_type) %>% 
   summarise(total_favourite_count = sum(favorite_count)) %>% 
   arrange(desc(total_favourite_count))
  

```



# Question 5.

Find the mean number of characters that a CodeClan tweet contains. You can either find a way to count the text in the text variable, or use the variable display_text_width which contains a count. Checking both would be a good idea if possible to see if they match (but it’s not essential).

```{r}
tweets %>% 
  mutate(text_width = str_length(text),
         .after = display_text_width) %>% 
  summarise(mean_characters = mean(text_width))


# tweets %>% 
#   summarise(mean_characters = mean(display_text_width))

# DISPLAY TEXT WIDTH IS LIES!!!!!!!!!!!!!!!!!!!!!!!!!!!!!

```


# Question 6.

The code_clan_info.csv data file contains status_url along with other info. Load this in, and join it to the code_clan_tweets tibble, so that you have a status_url for each tweet. Decide which variable to join the tibbles on.

```{r}
info <- read_csv("data/code_clan_info.csv")
```

```{r}
tweet_info <- left_join(tweets, info, by = "tweet_id")
tweet_info
```



# Question 7.

From your new joined data, create a new tibble codeclan_hashtags containing only tweet_id and hashtags, then convert hashtags into lowercase for analysis. Keep only those tweets with hashtags.

```{r}

codeclan_hashtags <- tweet_info %>% 
  drop_na(hashtags) %>% 
  select(tweet_id, hashtags) %>% 
  mutate(hashtags = str_to_lower(hashtags))
codeclan_hashtags

```

#4 Extensions

# Question 8.

Some tweets have more than one hashtag, denoted by the c( symbols at the start of the string. Use the str_sub() and str_detect() functions and your codeclan_hashtags tibble from above to find all the cases in which the hashtag string begins with charactersc(.


```{r}
# using str_detect
codeclan_hashtags %>% 
  filter((str_detect(hashtags, "^c\\(")) == TRUE)
  
  
  # using str_sub       
codeclan_hashtags %>%
  filter(if_else((str_sub(hashtags, 1, 2)) == "c(", TRUE, FALSE) == TRUE) 

# I wasnt clear if this was to be 2 methods, i.e. 1 using each function or not
# Thats the way i've answered it anyway
```

# Question 9.

Use the str_detect() function to find all tweets with text that mentions “Edinburgh”, and count how many such tweets there are.


```{r}
tweets %>% 
  # mutate(text = str_to_lower(text)) - Don't need this just use regex with (?i)
  filter(str_detect(text, "(?i)edinburgh") == TRUE) %>% 
  nrow()

# there are 33
```

# Question 10.

Use str_extract_all() with a regular expression to find out which Twitter users CodeClan have been tweeting.


```{r}

# The only punctuation allowed in a twitter name is an _ thanks google
tweets %>% 
  pull(text) %>% 
  str_extract_all("@[a-zA-Z0-9_]+") %>% 
  flatten_chr() # finds 256 twitter names

# Check we've found them all by counting just "@" symbols
# tweets %>%
#   pull(text) %>%
#   str_extract_all("@") %>%
#   flatten_chr() %>% 
#   nzchar() %>% 
#   sum()
# Theres definitely an easier way to do this but I'm sleepy and it works


```

